---
title: "DGE analysis"
output:
  html_notebook: default
  pdf_document: default
---
Load library and set working directory
```{r}
setwd("C:/Users/pradd/Desktop/Research Project/Control_new")
library(DESeq2)
library(ggplot2)
library(pheatmap)
library(EnhancedVolcano)
packageVersion("DESeq2")
```
```{r}
setwd("C:/Users/pradd/Desktop/Research Project/Control_new")
Counts <- read.delim("counts_table.csv", header= TRUE, row.names = 1, sep = ",")
Counts
```
Number of Gene-IDs inside the data:
```{r}
nrow(Counts)
```
Get the NA's positons in the data:
```{r}
print("Row and Col positions of NA values")
which(is.na(Counts))
```
Creation of Counts table:
```{r}
Counts <- Counts[which(rowSums(Counts) > 50),]
Counts
condition <- factor(c("C","C","C","C","C","C","C","C","T","T","T","T","T","T","T","T"))
coldata <- data.frame(row.names = colnames(Counts), condition)
coldata
```
DGE
```{r}
dds <- DESeqDataSetFromMatrix(countData = Counts, colData = coldata, design = ~condition)
dds <- DESeq(dds)
```
Save normalized read Counts:
```{r}
normalized_counts <- counts(dds, normalized = TRUE)
head(normalized_counts)
write.csv(normalized_counts, 'normalized_counts.csv')
```
```{r}
res <- results(dds,contrast = c('condition', 'T', 'C'), alpha = 0.05)
summary(res)
```
MA Plot
```{r}
plotMA(res, cex=0.7, ylim=c(-10,10))
abline(h=c(-1,1), col="red", lwd=3)
resultsNames(dds)
```
```{r}
res2 <- results(dds,contrast = c('condition', 'C', 'T'), alpha = 0.05)
#MA plot control vs Treatment
plotMA(res2, cex=0.7, ylim=c(-10,10))
abline(h=c(-1,1), col="red", lwd=3)
resultsNames(dds)
```
Shrinkage of effect size
```{r}
resLFC <- lfcShrink(dds,coef = "condition_T_vs_C", type="apeglm")
plotMA(resLFC, cex=0.7, ylim=c(-10,10))
abline(h=c(-1,1), col="red", lwd=3)
resultsNames(dds)
```
Dispersion Plot
```{r}
plotDispEsts(dds, main="Dispersion Plot")
```
PCA plot
```{r}
rld <- rlogTransformation(dds,blind = FALSE)
head(assay(rld))
hist(assay(rld))
PCAA<- plotPCA(rld, intgroup='condition')
PCAA +ggtitle('PCA Plot')
#+geom_text(aes(label=name),size=5)
```
Enhanced Volcano Plot
```{r fig.height=10}
EnhancedVolcano(res,
                lab=rownames(res),
                x='log2FoldChange',
                y='pvalue',
                title = 'Treatment vs. Control',
                pCutoff = 10e-4,
                FCcutoff= 0.5,
                pointSize = 3.0,
                labSize = 5.0)
```
Heatmap generation

SampleDists <- dist(+(assay(rld)))
library("RColorBrewer")
SampleDistMatrix <- as.matrix(SampleDists)
colnames(SampleDistMatrix)
colors<- colorRampPalette(rev(brewer.pal(9,"Blues")) )(255)
pheatmap(SampleDistMatrix, 
         clustering_distance_rows = SampleDists,
         clustering_distance_cols = SampleDists,
         col=colors)
         
top-genes<- res[order(res$padj), ][1:10] 
class(top-genes) 
top-genes         

```{r}
heatmap(assay(dds), scale = "row", dendrogram="row")
```
```

```
```{r}
setwd("C:/Users/pradd/Desktop/Research Project/Control_new")
write.csv(res, file="Deseq3C result")
```









